\TitlePt

### 2016-03-28 paper-fatiando.tex
Define API the first time it appears
Fixed #2
Add paragraph and notes about the current version
Also highlight that it was developed by me.
Suggestions by Val

This chapter was published in the Proceedings of the 12th Python in Science
Conference (Scipy 2013).
\url{http://conference.scipy.org/proceedings/scipy2013/uieda.html}

@@

Geophysics is the science of using physical observations of the Earth to infer
its inner structure. Generally, this is done with a variety of numerical
modeling techniques and inverse problems. The development of new algorithms
usually involves copy and pasting of code, which leads to errors and poor code
reuse. Fatiando a Terra is a Python library that aims to automate common tasks
and unify the modeling pipeline inside of the Python language. This allows
users to replace the traditional shell scripting with more versatile and
powerful Python scripting. The library can also be used as an API for
developing stand-alone programs.  Algorithms implemented in Fatiando a Terra
can be combined to build upon existing functionality. This flexibility
facilitates prototyping of new algorithms and quickly building interactive
teaching exercises. In the future, we plan to continuously implement sample
problems to help teach geophysics as well as classic and state-of-the-art
algorithms.

@@

The seismic method is among the most widely studied due to the high industry
demand. Thus, a range of well established open-source software have been
developed for seismic processing. These include Seismic Un*x \citep[][
\url{http://www.cwp.mines.edu/cwpcodes/}]{stockwelljr.1999}, Madagascar
\citep[][ \url{http://www.ahay.org/}]{madagascardevelopmentteam2013}, OpendTect
(\url{http://opendtect.org}), and GêBR (\url{http://www.gebrproject.com}). A
noteworthy open-source project that is not seismic related is the Generic
Mapping Tools (GMT) project \citep[][
\url{http://gmt.soest.hawaii.edu/}]{wessel1991}. The GMT are a well established
collection of command-line programs for plotting maps with a variety of
different map projections.  For geodynamic modeling there is the Computational
Infrastructure for Geodynamics (\url{http://www.geodynamics.org}), which has
grouped various well documented software packages. However, even with this wide
range of well maintained software projects, many geophysical modeling software
that are provided online still have no open-source license statement, have
cryptic I/O files, are hard to integrate into a pipeline, and make code reuse
and remixing challenging. Some of these problems are being worked on by the
Solid Earth Teaching and Research Environment (SEATREE) \citep[][
\url{http://geosys.usc.edu/projects/seatree/}]{milner2009} by providing a
common graphical interface for previously existing software. The numerical
computations are performed by the pre-existing underlying C/Fortran programs.
Conversely, the SEATREE code (written in Python) handles the I/O and user
interface. This makes the use of these tools easier and more approachable to
students. However, the lack of a common API means that the code for these
programs cannot be easily combined to create new modeling tools.

Fatiando a Terra (\url{http://www.fatiando.org}) aims at providing such an API
for geophysical modeling. Functions in the \texttt{fatiando} package use
compatible data and mesh formats so that the output of one modeling function
can be used as input for another. Furthermore, routines can be combined and
reused to create new modeling algorithms.  Fatiando a Terra also automates
common tasks such as griding, map plotting with Matplotlib \citep[][
\url{http://matplotlib.org}]{hunter2007}, and 3D plotting with Mayavi \citep[][
\url{http://code.enthought.com/projects/mayavi}]{ramachandran2011}.  Version
0.1 of Fatiando a Terra is focused on gravity and magnetic methods because this
is the main focus of the developers.  However, simple ``toy'' problems for
seismology and geothermics are available and can be useful for teaching
geophysics.

@@

The Fatiando a Terra package provides an API to develop modeling
algorithms for a variety of geophysical methods. The current version
(0.1) has a few state-of-the-art gravity and magnetic modeling and
inversion algorithms. There are also toy problems in gravity, seismics
and seismology that are useful for teaching basic concepts of
geophysics, modeling, and inverse problems.

>>>>

This chapter was published in the Proceedings of the 12th Python in Science
Conference in 2013
(\url{http://conference.scipy.org/proceedings/scipy2013/uieda.html}).
It describes the open-source software library Fatiando a Terra and the state of
the project at the time of the conference.
Fatiando a Terra was developed by me as part of my PhD thesis work.
The library has grown since 2013 and now contains more features than described
here.
For example, the algorithm for gravitational forward modeling with tesseroids
presented in Chapter~\ref{chap:tesseroids} is also implemented in Fatiando a
Terra.
Furthermore, the gravity inversion method presented in Chapter~\ref{chap:moho}
is implemented using functions and classes from Fatiando a Terra and will be
included in a future release of the library.
The official website (\url{http://www.fatiando.org}) contains more up-to-date
information about the project.

@@

Geophysics is the science of using physical observations of the Earth to infer
its inner structure. Generally, this is done with a variety of numerical
modeling techniques and inverse problems. The development of new algorithms
usually involves copy and pasting of code, which leads to errors and poor code
reuse. Fatiando a Terra is a Python library that aims to automate common tasks
and unify the modeling pipeline inside of the Python language. This allows
users to replace the traditional shell scripting with more versatile and
powerful Python scripting. The library can also be used as an API (Application
Programming Interface) for developing stand-alone programs.  Algorithms
implemented in Fatiando a Terra can be combined to build upon existing
functionality. This flexibility facilitates prototyping of new algorithms and
quickly building interactive teaching exercises. In the future, we plan to
continuously implement sample problems to help teach geophysics as well as
classic and state-of-the-art algorithms.

@@

The seismic method is among the most widely studied due to the high industry
demand. Thus, a range of well established open-source software have been
developed for seismic processing. These include Seismic Un*x \citep[][
\url{http://www.cwp.mines.edu/cwpcodes/}]{stockwelljr.1999}, Madagascar
\citep[][ \url{http://www.ahay.org/}]{madagascardevelopmentteam2013}, OpendTect
(\url{http://opendtect.org}), and GêBR (\url{http://www.gebrproject.com}). A
noteworthy open-source project that is not seismic related is the Generic
Mapping Tools (GMT) project \citep[][
\url{http://gmt.soest.hawaii.edu/}]{wessel1991}. The GMT are a well established
collection of command-line programs for plotting maps with a variety of
different map projections.  For geodynamic modeling there is the Computational
Infrastructure for Geodynamics (\url{http://www.geodynamics.org}), which has
grouped various well documented software packages. However, even with this wide
range of well maintained software projects, many geophysical modeling software
that are provided online still have no open-source license statement, have
cryptic I/O files, are hard to integrate into a pipeline, and make code reuse
and remixing challenging. Some of these problems are being worked on by the
Solid Earth Teaching and Research Environment (SEATREE) \citep[][
\url{http://geosys.usc.edu/projects/seatree/}]{milner2009} by providing a
common graphical interface for previously existing software. The numerical
computations are performed by the pre-existing underlying C/Fortran programs.
Conversely, the SEATREE code (written in Python) handles the I/O and user
interface. This makes the use of these tools easier and more approachable to
students. However, the lack of a common Application Programming Interface (API)
means that the code for these programs cannot be easily combined to create new
modeling tools.

Fatiando a Terra (\url{http://www.fatiando.org}) aims at providing such an API
for geophysical modeling. Functions in the \texttt{fatiando} package use
compatible data and mesh formats so that the output of one modeling function
can be used as input for another. Furthermore, routines can be combined and
reused to create new modeling algorithms.  Fatiando a Terra also automates
common tasks such as griding, map plotting with Matplotlib \citep[][
\url{http://matplotlib.org}]{hunter2007}, and 3D plotting with Mayavi \citep[][
\url{http://code.enthought.com/projects/mayavi}]{ramachandran2011}.  Version
0.1\footnote{
    This was the current version in 2013 when this chapter was published in the
    conference proceedings. As of April 2016, the latest version is 0.3.}
of Fatiando a Terra is focused on gravity and magnetic methods because this is
the main focus of the developers.  However, simple ``toy'' problems for
seismology and geothermics are available and can be useful for teaching
geophysics.

@@

The Fatiando a Terra package provides an API to develop modeling
algorithms for a variety of geophysical methods. The current version
(0.1)\footnote{
    As of April 2016, the latest version is 0.3.}
has a few state-of-the-art gravity and magnetic modeling and
inversion algorithms. There are also toy problems in gravity, seismics
and seismology that are useful for teaching basic concepts of
geophysics, modeling, and inverse problems.
